\ttitle

### 2015-08-26 Chapters/Pre_Post_SIRS.tex
Pre_Post_SIRS - Paper review complete
- Look at the number of bloody mistakes

An elevated preoperative systemic inflammatory response was associated with increased complications after colorectal surgery \parencite{moyes_preoperative_2009, kubo_elevated_2013}, oesophagectomy \parencite{vashist_glasgow_2010} as well as liver surgery for colorectal metastases \parencite{neal_preoperative_2011}. 
The modified Glasgow Prognostic Score in particular, which uses the combination of C-reactive protein and serum albumin, has been reported to be associated with increased incidence of complications \parencite{moyes_preoperative_2009, mohri_correlation_2014, vashist_glasgow_2010}.
Elevated preoperative CRP levels have also been reported to be associated with increased incidence of complications including infections and renal dysfunction as well as increased in-hospital mortality after cardiac surgery \parencite{lorenzo_increased_2012, mezzomo_preoperative_2011, kim_predictive_2009, biancari_preoperative_2003, boeken_increased_1998}.

Moreover, an elevated postoperative systemic inflammatory response in the first few days after surgery is associated with increased incidence of infective complications after a wide variety of thoraco-abdominal procedures \parencite{singh_systematic_2014, platt_c-reactive_2012, dutta_persistent_2011, welsch_persisting_2008} as well as other types of surgery \parencite{mcneer_early_2010, laporta_baez_c-reactive_2011}.
%Include some recent reviews pp singh, bjs as well as check Donny's email.
The magnitude of this postoperative inflammatory response has also been reported to be associated with the severity of the complications \parencite{mcsorley_postoperative_2015}.

In patients undergoing pancreaticoduodenectomy, preoperative systemic inflammation may be affected by several factors. 
These include the presence of obstructive jaundice with or without cholangitis, preoperative biliary intervention including endoscopic retrograde cholangio-pancreatography for diagnosis or biliary drainage and in some patients acute or chronic pancreatitis either due to obstruction of the main pancreatic duct or due to other causes. 
The effect this `priming' of the immune system on postoperative outcomes is poorly understood in this cohort of patients.

Chronic inflammation is a recognised feature of obesity which is increasingly common in patients undergoing major surgery for pancreatic and other gastro-intestinal cancers. 
The impact of obesity, especially visceral obesity, on complications after pancreaticoduodenectomy remains controversial with some authors reporting that obesity was associated with increased incidence of complications \parencite{house_preoperative_2008, ramsey_body_2011} while others reporting similar outcomes in obese and non-obese patients \parencite{khan_does_2010, tsai_impact_2010, balentine_obesity_2011}.

More recently, levels of adipocytokines, inflammatory mediators produced exclusively in adipose tissue, have been reported to be associated with postoperative surgical site infections after colorectal \parencite{ortega-deballon_preoperative_2013, matsuda_preoperative_2009} and gastric cancer surgery \parencite{yamamoto_association_2013}.
Sarcopenia has been reported to be associated with  elevated postoperative systemic inflammation after colorectal surgery \parencite{reisinger_sarcopenia_2015}.
This emphasises the importance of adipose tissue metabolism and body composition in the preoperative systemic inflammatory status of surgical patients.

@@

The modified Glasgow Prognostic Score (mGPS) was calculated as shown in Table \ref{table:mGPS} on p\pageref{table:mGPS} in Chapter \ref{ch_intro}. 
The neutrophil-lymphocyte ratio was calculated by dividing the preoperative neutrophil count by the preoperative lymphocyte count and a threshold of 5.0 was used to dichotomise this variable. 
Standard thresholds were used to categorise other biochemical parameters. 
Obstructive jaundice was defined as serum bilirubin $>$35 $\mu$mol/L while severe obstructive jaundice was defined as serum bilirubin $>$250 $\mu$mol/L.

\subsection{BMI and body composition}
Body Mass Index (BMI) was categorised using the World Health Organisation thresholds as shown in Table \ref{table:bmi_who} on p\ref{table:bmi_who} in Chapter \ref{ch_intro}. 
Only one patient had a BMI less than 18.5 kg/m$2$ and was excluded from analysis involving BMI.
Body composition was calculated in a subset of these patients using preoperative computed tomography of the abdomen. 
The methodology used in the calculation of the individual components of body composition including visceral fat, subcutaneous fat and skeletal muscle is described in detail in Section \ref{sec:bodycomp_calculation}. 
Continuous data were converted into categorical data using tertiles.

\subsection{Comorbidity and CPET}
The Scottish Index of Multiple Deprivation (SIMD) Quintile Scores were calculated from the post-code of the patient's primary residence and dichotomised into two groups with scores of 1-3 and 4-5 respectively.
The POSSUM Physiology Score was calculated based on 11 physiological parameters (cardiac disease including hypertension, ischaemic heart disease and heart failure, respiratory disease causing breathlessness on exertion and COPD, ECG changes, pulse rate, blood pressure, haemoglobin, white cell count, serum sodium, serum potassium, serum urea and Glasgow Coma Scale) as described in Table \ref{table:intro_possum} on p\pageref{table:intro_possum}.

In patients who underwent cardiopulmonary exercise testing, $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak were compared against postoperative systemic inflammation. 
$\dot{V}_{O_2}$AT was dichotomised using a value of 10 ml/kg/min while $\dot{V}_{O_2}$Peak was dichotomised using a value of 16 ml/kg/min. 
Cardiopulmonary exercise testing methodology has been described in Section \ref{sec:cpx_method}.

@@

Pancreaticoduodenectomy was performed in 188 patients (126 male, 67\%) during the study period.
Preoperative C-reactive protein was elevated in 70 (37.6\%) patients while just over half the patients had a low preoperative serum albumin (96, 51.1\%). 
The modified Glasgow Prognostic Score revealed normal preoperative systemic inflammatory status (mGPS = 0) in 116 (62.4\%) patients, mildly systemic inflammation in 16 (8.6\%) and severe inflammation in 54 (29.0\%) patients. 
Obstructive jaundice was present in 44 (23.4\%) patients and severe obstructive jaundice was present in 31 (16.5\%) of patients. 
BMI data was available in 167 patients. 
More than half of these patients were overweight or obese with a BMI $>$ 25 in 87 (52.1\%) patients.
All the variables necessary for calculation of the POSSUM Physiology Score, a composite score of comorbidity and preoperative biochemistry, were available in 180 patients and it was elevated in 87 (48.3\%) patients.

@@

Preoperative hypoalbuminemia on the other hand was associated with lower median postoperative CRP levels starting with POD 3. 
The median difference in postoperative CRP increased from 30 mg/l on POD 3 (p$<$0.045) to 53 mg/l by POD 6 (p=0.002) with patients with hypoalbuminemia having a lower CRP. 
This relationship is shown in Fig. \ref{fig:sirs_alb_crp}.
Preoperative hypoalbuminemia persisted postoperatively throughout the first week (p$<$0.001, Fig. \ref{fig:sirs_alb_alb}).
There was no relationship between preoperative serum albumin levels and postoperative neutrophil count (Fig. \ref{fig:sirs_alb_neut}).

The trend in postoperative CRP appeared to differ depending on whether preoperative CRP and albumin were independently deranged or not.
Fig. \ref{fig:sirs_crp_crp_alb} and Table \ref{table:sirs_crp_with_alb} show the results of this analysis.
When both preoperative CRP and preoperative albumin were normal, the trend in postoperative CRP is shown by the solid line (Line 0 in Fig. \ref{fig:sirs_crp_crp_alb}). 
When preoperative CRP was elevated in the presence of a normal preoperative albumin, postoperative CRP levels were significantly elevated and remained elevated during the first postoperative week (Line 1 in Fig. \ref{fig:sirs_crp_crp_alb}). 
However, hypoalbuminemia was associated with lower postoperative CRP levels regardless of whether preoperative CRP was normal (Line 2 in Fig. \ref{fig:sirs_crp_crp_alb}) or elevated (Line 3 in Fig. \ref{fig:sirs_crp_crp_alb}).

\subsection{Obstructive jaundice vs. postoperative inflammation}
Obstructive jaundice was associated with significant differences in the postoperative systemic inflammatory response. 
Postoperative serum albumin levels were lower in jaundiced patients and with the severely jaundiced patients having the lowest levels (p$<$0.001, Fig. \ref{fig:sirs_bil_alb}). 
There were no significant changes in the trends of albumin levels and it is likely that preoperative hypoalbuminemia persisted postoperatively in these patients.

However, the trends in postoperative CRP were significantly different between the jaundiced and non-jaundiced patients. 
Obstructive jaundice was associated with a significantly lower peak CRP on POD 2 (p=0.001) and this persisted until POD 6 (p=0.005, Table \ref{table:sirs_crp}, \ref{table:sirs_crp_pvalues}).
Median postoperative CRP had an inverse relationship with severity of preoperative obstructive jaundice (Fig. \ref{fig:sirs_bil_crp}).
Obstructive jaundice was also associated with a delayed rise in the neutrophil count on POD 6 (p$<$0.05) and POD 7 (p$<$0.05) although this relationship was less significant (Fig. \ref{fig:sirs_bil_neut}).

@@

There was no significant, persistent relationship between postoperative CRP and neutrophil count and preoperative neutrophil-lymphocyte ratio, SIMD score, body mass index, preoperative haemoglobin levels or the POSSUM Physiology score. 
However, postoperative hypoalbuminemia was associated with all the preoperative characteristics studied with the exception of BMI and SIMD (Table \ref{table:sirs_alb}, \ref{table:sirs_alb_pvalues}).

\subsection{Body composition vs. postoperative inflammation}
Body composition data was available for 90 patients.
Visceral adipose tissue area (cm$^2$) was divided into tertiles as low (n=30, median 57.1 cm$^2$, IQR 29.0-62.4), moderate (n=30, 128.9 cm$^2$, 119.0-150.8)and high (n=30, 241.2 cm$^2$, 221.9-313.1). 
Subcutaneous adipose tissue area was divided into tertiles as low (n=30, median 94.1 cm$^2$, IQR 63.2-102.5), moderate (n=30, median 152.5 cm$^2$, IQR 141.3-175.3) and high (n=30, median 253.9 cm$^2$, IQR 203.2-290.3).
Skeletal muscle area was also divided into tertiles as low (n=30, median 96.7 cm$^2$, IQR 91.9-102.5), moderate (n=30, median 121.7 cm$^2$, IQR 116.4-130.9) and high (n=30, median 153.0 cm$^2$, IQR 145.3-167.6).

The relationship between body composition and postoperative CRP, albumin and neutrophil count is shown in Tables \ref{table:sirs_bodycomp_crp}, \ref{table:sirs_bodycomp_alb} and \ref{table:sirs_bodycomp_neut} respectively.
Postoperative CRP and neutrophil count were not related to any of the components of body composition aside from a weak linear association between neutrophil count on POD 2 and visceral fat area (p=0.046).
However, lower preoperative skeletal muscle area was associated with lower postoperative serum albumin on days 1 to 5 (all p<0.05).

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_crp_neut}
		\caption{Preop. CRP vs. postop. Neutrophil Count}
		\label{fig:sirs_crp_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_alb_neut}
		\caption{Preop. Albumin vs. postop. Neutrophil Count}
		\label{fig:sirs_alb_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_bil_neut}
		\caption{Preop. Bilirubin vs. postop. Neutrophil Count}
		\label{fig:sirs_bil_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_at_neut}
		\caption{Preop. $\dot{V}_{O_2}$AT vs. postop. Neutrophil Count}
		\label{fig:sirs_at_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

The role of postoperative CRP in predicting complications after surgery has been well established. 
A recent meta-analysis of six studies involving 1832 patients reported that CRP level less than 135 mg/l on the fourth postoperative day had a high negative predictive value of 89\% for infectious complications \parencite{warschkow_safe_2012}. 
Postoperative CRP has been shown to predict complications after pancreaticoduodenectomy by us (\ref{ch_crp_comp}) as well as other authors \parencite{welsch_persisting_2008, hiyoshi_usefulness_2013, kosaka_multivariate_2014}.
The role of preoperative systemic inflammation as measured by the modified Glasgow Prognostic Score or the neutrophil-lymphocyte ratio has also been reported to be adversely affect long-term survival in patients undergoing potentially curative surgery in over a hundred studies involving a variety of gastrointestinal and non-gastrointestinal cancers.

@@

We hypothesized that preoperative systemic inflammation may adversely affect postoperative immuno-modulation resulting in an abnormal response in patients whose immune systems have been `primed' before the surgical insult. 
This may affect the healing processes as well as the patient's ability to clear endotoxins and may predispose the patient to infective and other complications.
Our results appear to support this hypothesis.

In this study, raised preoperative CRP in the absence of hypoalbuminemia was associated with elevated postoperative systemic inflammation.
However, preoperative hypoalbuminemia was associated with significantly lower levels of postoperative CRP.
These apparently opposite effects of preoperative CRP and albumin on the postoperative systemic inflammatory response is further evidenced by the fact that the effect of preoperative CRP on postoperative inflammation is attenuated by the presence of preoperative hypoalbuminemia.
These results appear to suggest when hypoalbuminemia is present as part of preoperative systemic inflammation, not only is the exaggerated postoperative response lost but the response appears to be dampened.

%Hypoalbuminemia
The impact of hypoalbuminemia on the acute phase response was noted by Christou and co-workers who performed a delayed type hypersensitivity (DTH) skin test in addition to measuring acute phase markers including CRP, albumin, white cell count, haemoglobin and immunoglobulins in 245 patients prior to undergoing gastrointestinal surgery \parencite{christou_estimating_1989}.
They noted that hypoalbuminemia and anergy to the DTH skin test were the only variables independently associated with postoperative sepsis-related mortality.
Moreover, both low albumin and elevated CRP were significantly associated with cutaneous anergy.
Cutaneous anergy is one of the hallmarks of the compensatory anti-inflammatory response (CARS) \parencite{ward_compensatory_2008}.

@@

%Obstructive jaundice
In patients with pancreatic cancer, malignant obstructive jaundice further complicates the preoperative inflammatory status and the behaviour of the immune system \parencite{nehez_compromise_2002}. 
We reported in Chapter \ref{ch_cpet_jaundice} that obstructive jaundice was associated with elevated preoperative CRP and low serum albumin on the day before surgery with a linear relationship with severity of jaundice (Table \ref{table:cpet_oj_bloods} on p\pageref{table:cpet_oj_bloods}).
Hypoalbuminemia in jaundiced patients persisted after surgery in the present study.
However, the CRP response appears to have reversed in jaundiced patients.
While preoperative CRP was elevated in jaundiced patients, CRP levels after surgery were significantly lower in jaundiced patients with a negative linear relationship with severity of jaundice.

@@

Immunosuppression and immune dysregulation have been recognised in patients with obstructive jaundice \parencite{scott-conner_pathophysiology_1994}.
Our results are similar to those reported by Mackenzie and Woodhouse who studied the CRP response to bacteraemia in 126 patients critically ill patients with or without liver dysfunction \parencite{mackenzie_c-reactive_2006}. 
Liver dysfunction was defined as serum bilirubin $>$ 20 $\mu$mol/l and prothrombin time longer than 18 seconds. 
The CRP response to bacteraemia was significantly lower in patients with liver dysfunction than in those with normal liver function (146 vs. 103 mg/l, p=0.03). 
These findings have been corroborated by other authors in patients with cirrhosis \parencite{pieri_c-reactive_2014, janum_c-reactive_2011}.

%There appears to be some evidence from animal studies that the immunosuppression due to obstructive jaundice may be a therapeutic target \parencite{padillo_anti-tnf-alpha_2007}.

%Aerobic capacity/comorbidity
Although CRP on POD 0 was significantly different between patients with a normal $\dot{V}_{O_2}$AT/$\dot{V}_{O_2}$Peak and those with low aerobic capacity, this difference was not present on other days. 
The lack of association between aerobic capacity, POSSUM Physiology Score and postoperative CRP in this study may have been due to the dominant effect of preoperative systemic inflammation and obstructive jaundice on the postoperative inflammatory response or simply due to any influence of aerobic capacity on postoperative inflammation.
However, the association of postoperative hypoalbuminemia with low aerobic capacity, POSSUM Physiology Score and anaemia appears to suggest a link between comorbidity and systemic inflammation and/or nutritional status.
Bernal and co-workers studied 399 patients undergoing liver transplantation who had cardiopulmonary exercise testing as part of their preoperative work-up and reported that hypoalbuminemia was associated with both low $\dot{V}_{O_2}$AT and low $\dot{V}_{O_2}$Peak \parencite{bernal_aerobic_2014}.

%Body Composition
The association between hypoalbuminemia and low aerobic capacity may be explained by lower skeletal muscle and greater subcutaneous fat in patients with persistent postoperative hypoalbuminemia.
Alternate explanations 
Richards and co-workers reported that elevated preoperative systemic inflammation was related to low skeletal muscle index \parencite{hassen_preoperative_2007}.
Hassen and co-workers reported that fat-free mass and skeletal muscle were inversely related to the severity of the postoperative inflammatory response following endo-vascular surgery for abdominal aortic aneurysm \parencite{hassen_preoperative_2007}.
Elevated systemic inflammation is related to both obesity as well as sarcopenia. 
`Sarcopenic-obesity' in patients with pancreatic cancer where there is relative loss of skeletal muscle with preservation of adipose tissue puts these patients at particular risk for immune dysfunction \parencite{berg_adipose_2005, reisinger_sarcopenia_2015}, post-pancreatectomy complications \parencite{joglekar_sarcopenia_2015} and shorter long-term survival in patients with pancreatic cancer \parencite{tan_sarcopenia_2009, peng_impact_2012}.

>>>>

Elevated preoperative systemic inflammation was associated with increased complications after colorectal surgery \parencite{moyes_preoperative_2009, kubo_elevated_2013}, oesophagectomy \parencite{vashist_glasgow_2010} as well as liver surgery for colorectal metastases \parencite{neal_preoperative_2011}. 
The modified Glasgow Prognostic Score in particular, which uses the combination of C-reactive protein and serum albumin, has been reported to be associated with increased incidence of complications \parencite{moyes_preoperative_2009, mohri_correlation_2014, vashist_glasgow_2010}.
Elevated preoperative CRP levels have also been reported to be associated with increased incidence of complications including infections and renal dysfunction as well as increased in-hospital mortality after cardiac surgery \parencite{lorenzo_increased_2012, mezzomo_preoperative_2011, kim_predictive_2009, biancari_preoperative_2003, boeken_increased_1998}.

Moreover, an elevated postoperative systemic inflammatory response in the first few days after surgery was associated with increased incidence of infective complications after a wide variety of thoraco-abdominal procedures \parencite{singh_systematic_2014, platt_c-reactive_2012, dutta_persistent_2011, welsch_persisting_2008} as well as other types of surgery \parencite{mcneer_early_2010, laporta_baez_c-reactive_2011}.
%Include some recent reviews pp singh, bjs as well as check Donny's email.
The magnitude of this postoperative inflammatory response has also been reported to be associated with the severity of the complications \parencite{mcsorley_postoperative_2015}.

In patients undergoing pancreaticoduodenectomy, preoperative systemic inflammation may be affected by several factors. 
These include the presence of obstructive jaundice with or without cholangitis, preoperative biliary intervention including endoscopic retrograde cholangio-pancreatography for diagnosis or biliary drainage and in some patients acute or chronic pancreatitis either due to obstruction of the main pancreatic duct or due to other causes. 
The effect of this `priming' of the immune system on postoperative outcomes is poorly understood in this cohort of patients.

Chronic inflammation is also a recognised feature of obesity which is increasingly common in patients undergoing major surgery for pancreatic and other gastro-intestinal cancers. 
The impact of obesity, especially visceral obesity, on complications after pancreaticoduodenectomy remains controversial with some authors reporting that obesity was associated with increased incidence of complications \parencite{house_preoperative_2008, ramsey_body_2011} while others reporting similar outcomes in obese and non-obese patients \parencite{khan_does_2010, tsai_impact_2010, balentine_obesity_2011}.

More recently, levels of adipocytokines, inflammatory mediators produced exclusively in adipose tissue, have been reported to be associated with postoperative surgical site infections after colorectal \parencite{ortega-deballon_preoperative_2013, matsuda_preoperative_2009} and gastric cancer surgery \parencite{yamamoto_association_2013}.
Sarcopenia has also been reported to be associated with  elevated postoperative systemic inflammation after colorectal surgery \parencite{reisinger_sarcopenia_2015}.
This emphasises the importance of adipose tissue metabolism and body composition in the preoperative systemic inflammatory status of surgical patients.

@@

The modified Glasgow Prognostic Score (mGPS) was calculated as shown in Table \ref{table:mGPS} on page \pageref{table:mGPS} in Chapter \ref{ch_intro}. 
The neutrophil-lymphocyte ratio was calculated by dividing the preoperative neutrophil count by the preoperative lymphocyte count and a threshold of 5.0 was used to dichotomise this variable. 
Standard thresholds were used to categorise other biochemical parameters. 
Obstructive jaundice was defined as serum bilirubin $>$35 $\mu$mol/L while severe obstructive jaundice was defined as serum bilirubin $>$250 $\mu$mol/L.

\subsection{BMI and body composition}
Body Mass Index (BMI) was categorised using the World Health Organisation thresholds as shown in Table \ref{table:bmi_who} on p \pageref{table:bmi_who} in Chapter \ref{ch_intro}. 
Only one patient had a BMI less than 18.5 kg/m$2$ and was excluded from analysis involving BMI.
Body composition was calculated in a subset of these patients using preoperative computed tomography of the abdomen. 
The methodology used in the calculation of the individual components of body composition including visceral fat, subcutaneous fat and skeletal muscle is described in detail in Section \ref{sec:bodycomp_calculation} on page \pageref{sec:bodycomp_calculation}. 
Continuous data were converted into categorical data using tertiles.

\subsection{Comorbidity and CPET}
The Scottish Index of Multiple Deprivation (SIMD) Quintile Scores were calculated from the post-code of the patient's primary residence and dichotomised into two groups with scores of 1-3 and 4-5 respectively.
Lower scores indicated greater deprivation.
The POSSUM Physiology Score was calculated based on 11 physiological parameters (cardiac disease including hypertension, ischaemic heart disease and heart failure, respiratory disease causing breathlessness on exertion and COPD, ECG changes, pulse rate, blood pressure, haemoglobin, white cell count, serum sodium, serum potassium, serum urea and Glasgow Coma Scale) as described in Table \ref{table:intro_possum} on page \pageref{table:intro_possum}.

In patients who underwent cardiopulmonary exercise testing, $\dot{V}_{O_2}$AT and $\dot{V}_{O_2}$Peak were compared against postoperative systemic inflammation. 
$\dot{V}_{O_2}$AT was dichotomised using a value of 10 ml/kg/min while $\dot{V}_{O_2}$Peak was dichotomised using a value of 16 ml/kg/min. 
Cardiopulmonary exercise testing methodology has been described in Section \ref{sec:cpx_method} on page \pageref{sec:cpx_method}.

@@

Pancreaticoduodenectomy was performed in 188 patients (126 male, 67\%) during the study period.
Preoperative C-reactive protein was elevated in 70 (37.6\%) patients while just over half the patients had a low preoperative serum albumin (96, 51.1\%). 
The modified Glasgow Prognostic Score revealed normal preoperative systemic inflammatory status (mGPS = 0) in 116 (62.4\%) patients, mild systemic inflammation in 16 (8.6\%) and severe inflammation in 54 (29.0\%) patients. 
Obstructive jaundice was present in 44 (23.4\%) patients and severe obstructive jaundice was present in 31 (16.5\%) patients. 
BMI data was available in 167 patients. 
More than half of these patients were overweight or obese with a BMI $>$ 25 in 87 (52.1\%) patients.
All the variables necessary for calculation of the POSSUM Physiology Score, a composite score of comorbidity and preoperative biochemistry, were available in 180 patients and it was elevated in 87 (48.3\%) patients.

@@

Preoperative hypoalbuminemia on the other hand was associated with lower median postoperative CRP levels starting with POD 3. 
The median difference in postoperative CRP between patients with or without preoperative hypoalbuminemia was 30 mg/l on POD 3 (p$<$0.045) and 53 mg/l on POD 6 (p=0.002) with lower CRP levels in patients with hypoalbuminemia.
This relationship is shown in Fig. \ref{fig:sirs_alb_crp}.
Preoperative hypoalbuminemia persisted postoperatively throughout the first week (p$<$0.001, Fig. \ref{fig:sirs_alb_alb}).
There was no relationship between preoperative serum albumin levels and postoperative neutrophil count (Fig. \ref{fig:sirs_alb_neut}).

The trend in postoperative CRP appeared to differ depending on whether preoperative CRP and albumin were independently deranged.
Fig. \ref{fig:sirs_crp_crp_alb} and Table \ref{table:sirs_crp_with_alb} on page \pageref{fig:sirs_crp_crp_alb} show the results of this analysis.
When both preoperative CRP and preoperative albumin were normal, the trend in postoperative CRP is shown by the solid line (Line 0 in Fig. \ref{fig:sirs_crp_crp_alb}). 
When preoperative CRP was elevated in the presence of a normal preoperative albumin, postoperative CRP levels were significantly elevated and remained elevated during the first postoperative week (Line 1 in Fig. \ref{fig:sirs_crp_crp_alb}). 
However, hypoalbuminemia was associated with lower postoperative CRP levels regardless of whether preoperative CRP was normal (Line 2 in Fig. \ref{fig:sirs_crp_crp_alb}) or elevated (Line 3 in Fig. \ref{fig:sirs_crp_crp_alb}).

\subsection{Obstructive jaundice vs. postoperative inflammation}
Obstructive jaundice was associated with significant differences in the postoperative systemic inflammatory response. 
Postoperative serum albumin levels were lower in jaundiced patients with severely jaundiced patients having the lowest levels (p$<$0.001, Fig. \ref{fig:sirs_bil_alb}). 
There were no significant changes in the trends of albumin levels and it is likely that preoperative hypoalbuminemia persisted postoperatively in these patients.

However, the trends in postoperative CRP were significantly different between the jaundiced and non-jaundiced patients. 
Obstructive jaundice was associated with a significantly lower peak CRP on POD 2 (p=0.001) and this persisted until POD 6 (p=0.005, Table \ref{table:sirs_crp}, \ref{table:sirs_crp_pvalues}).
Median postoperative CRP had an inverse relationship with severity of preoperative obstructive jaundice (Fig. \ref{fig:sirs_bil_crp}).
Obstructive jaundice was associated with a delayed rise in the neutrophil count on POD 6 (p$<$0.05) and POD 7 (p$<$0.05) although this relationship was less significant (Fig. \ref{fig:sirs_bil_neut}).

@@

There was no significant, persistent relationship between postoperative CRP and neutrophil count and preoperative neutrophil-lymphocyte ratio, SIMD score, body mass index, preoperative haemoglobin levels or the POSSUM Physiology score. 
However, postoperative hypoalbuminemia was associated with all the preoperative characteristics studied with the exception of BMI and SIMD (Tables \ref{table:sirs_alb} and \ref{table:sirs_alb_pvalues}).

\subsection{Body composition vs. postoperative inflammation}
Body composition data was available for 90 patients.
Visceral adipose tissue area (cm$^2$) was divided into tertiles as low (n=30, median 57.1 cm$^2$, IQR 29.0-62.4), moderate (n=30, 128.9 cm$^2$, 119.0-150.8) and high (n=30, 241.2 cm$^2$, 221.9-313.1). 
Subcutaneous adipose tissue area was divided into tertiles as low (n=30, median 94.1 cm$^2$, IQR 63.2-102.5), moderate (n=30, median 152.5 cm$^2$, IQR 141.3-175.3) and high (n=30, median 253.9 cm$^2$, IQR 203.2-290.3).
Skeletal muscle area was also divided into tertiles as low (n=30, median 96.7 cm$^2$, IQR 91.9-102.5), moderate (n=30, median 121.7 cm$^2$, IQR 116.4-130.9) and high (n=30, median 153.0 cm$^2$, IQR 145.3-167.6).

The relationship between body composition and postoperative CRP, albumin and neutrophil count is shown in Tables \ref{table:sirs_bodycomp_crp}, \ref{table:sirs_bodycomp_alb} and \ref{table:sirs_bodycomp_neut} respectively.
Postoperative CRP and neutrophil count were not related to any of the components of body composition aside from a weak linear association between neutrophil count on POD 2 and visceral fat area (p=0.046).
However, lower preoperative skeletal muscle area was associated with lower postoperative serum albumin on days 1 to 5 (all p$<$0.05).

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_crp_neut}
		\caption{Preop. CRP vs. postop. Neutrophil count}
		\label{fig:sirs_crp_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_alb_neut}
		\caption{Preop. Albumin vs. postop. Neutrophil count}
		\label{fig:sirs_alb_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_bil_neut}
		\caption{Preop. Bilirubin vs. postop. Neutrophil count}
		\label{fig:sirs_bil_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

	\begin{subfigure}{0.48\textwidth}
		\centering
		\includegraphics[width=\textwidth]{Figures/sirs_at_neut}
		\caption{Preop. $\dot{V}_{O_2}$AT vs. postop. Neutrophil count}
		\label{fig:sirs_at_neut}
	\end{subfigure}	
\end{figure}
%==============================================================================

@@

The role of postoperative CRP in predicting complications after surgery has been well established. 
A recent meta-analysis of six studies involving 1832 patients reported that CRP level less than 135 mg/l on the fourth postoperative day had a high negative predictive value of 89\% for infectious complications \parencite{warschkow_safe_2012}. 
Postoperative CRP has been shown to predict complications after pancreaticoduodenectomy by us (chapter \ref{ch_crp_comp}) as well as other authors \parencite{welsch_persisting_2008, hiyoshi_usefulness_2013, kosaka_multivariate_2014}.
The role of preoperative systemic inflammation as measured by the modified Glasgow Prognostic Score or the neutrophil-lymphocyte ratio has also been reported to be adversely affect long-term survival in patients undergoing potentially curative surgery in over a hundred studies involving a variety of gastrointestinal and non-gastrointestinal cancers.

@@

We hypothesized that preoperative systemic inflammation may adversely affect postoperative immuno-modulation resulting in an abnormal response in patients whose immune systems have been `primed' before the surgical insult. 
This may affect the healing processes as well as the patient's ability to clear endotoxins and may predispose the patient to infective and other complications.
The results of the present study appear to support this hypothesis.

In this study, raised preoperative CRP in the absence of hypoalbuminemia was associated with elevated postoperative systemic inflammation.
However, preoperative hypoalbuminemia was associated with significantly lower levels of postoperative CRP.
These apparently opposite effects of preoperative CRP and albumin on the postoperative systemic inflammatory response is further evidenced by the fact that the effect of elevated preoperative CRP on postoperative inflammation is attenuated by the presence of preoperative hypoalbuminemia.
These results appear to suggest when hypoalbuminemia is present as part of preoperative systemic inflammation, not only is the exaggerated postoperative response due to elevated CRP lost but the response appears to be dampened.

%Hypoalbuminemia
The impact of hypoalbuminemia on the acute phase response was noted by Christou and co-workers who performed a delayed type hypersensitivity (DTH) skin test in addition to measuring acute phase markers including CRP, albumin, white cell count, haemoglobin and immunoglobulins in 245 patients prior to undergoing gastrointestinal surgery \parencite{christou_estimating_1989}.
They noted that hypoalbuminemia and anergy to the DTH skin test were the only variables independently associated with postoperative sepsis-related mortality.
Moreover, both low albumin and elevated CRP were significantly associated with cutaneous anergy.
Cutaneous anergy is one of the hallmarks of the compensatory anti-inflammatory response syndrome (CARS) \parencite{ward_compensatory_2008}.

@@

%Obstructive jaundice
In patients with pancreatic cancer, malignant obstructive jaundice further complicates the preoperative inflammatory status and the behaviour of the immune system \parencite{nehez_compromise_2002}. 
We reported in Chapter \ref{ch_cpet_jaundice} that obstructive jaundice was associated with elevated preoperative CRP and low serum albumin on the day before surgery with a linear relationship with severity of jaundice (Table \ref{table:cpet_oj_bloods} on page \pageref{table:cpet_oj_bloods}).
Hypoalbuminemia in jaundiced patients persisted after surgery in the present study.
However, the CRP response appears to have reversed in jaundiced patients.
While preoperative CRP was elevated in jaundiced patients, CRP levels after surgery were significantly lower in jaundiced patients with an inverse relationship with severity of jaundice.

@@

Immunosuppression and immune dysregulation have been recognised in patients with obstructive jaundice \parencite{scott-conner_pathophysiology_1994}.
Our results are similar to those reported by Mackenzie and Woodhouse who studied the CRP response to bacteraemia in 126 critically ill patients with or without liver dysfunction \parencite{mackenzie_c-reactive_2006}. 
Liver dysfunction was defined as serum bilirubin $>$ 20 $\mu$mol/l and prothrombin time longer than 18 seconds. 
The CRP response to bacteraemia was significantly lower in patients with liver dysfunction than in those with normal liver function (103 vs. 146 mg/l, p=0.03). 
These findings have been corroborated by other authors in patients with cirrhosis \parencite{pieri_c-reactive_2014, janum_c-reactive_2011}.

%Aerobic capacity/comorbidity
Although CRP on POD 0 was significantly different between patients with a normal $\dot{V}_{O_2}$AT/$\dot{V}_{O_2}$Peak and those with low aerobic capacity, this difference was not present on other days. 
The lack of association between aerobic capacity, POSSUM Physiology Score and postoperative CRP in this study may have been due to the dominant effect of preoperative systemic inflammation and obstructive jaundice on the postoperative inflammatory response.
However, the association of postoperative hypoalbuminemia with low aerobic capacity, POSSUM Physiology Score and anaemia appears to suggest a link between comorbidity and systemic inflammation and/or nutritional status.
Bernal and co-workers studied 399 patients undergoing liver transplantation who had cardiopulmonary exercise testing as part of their preoperative work-up and reported that hypoalbuminemia was associated with both low $\dot{V}_{O_2}$AT and low $\dot{V}_{O_2}$Peak \parencite{bernal_aerobic_2014}.

%Body Composition
The association between hypoalbuminemia and low aerobic capacity may be explained by lower skeletal muscle and greater subcutaneous fat in patients with persistent postoperative hypoalbuminemia.
Richards and co-workers reported that elevated preoperative systemic inflammation was related to low skeletal muscle index \parencite{richards_relationships_2012}.
Hassen and co-workers reported that fat-free mass and skeletal muscle were inversely related to the severity of the postoperative inflammatory response following endo-vascular surgery for abdominal aortic aneurysm \parencite{hassen_preoperative_2007}.

Elevated systemic inflammation is related to obesity as well as sarcopenia. 
`Sarcopenic-obesity' in patients with pancreatic cancer where there is relative loss of skeletal muscle with preservation of adipose tissue puts these patients at particular risk for immune dysfunction \parencite{berg_adipose_2005, reisinger_sarcopenia_2015}, post-pancreatectomy complications \parencite{joglekar_sarcopenia_2015} and shorter long-term survival \parencite{tan_sarcopenia_2009, peng_impact_2012}.
